# Frequentist Equivalence Testing {#Frequentist-Equivalence}

In the lectures, we mentioned equivalence testing is another option for statistically supporting there is no meaningful effect present. No technique can support the null hypothesis is precisely 0, so equivalence testing flips the standard hypothesis testing logic around. 

Instead of testing against a single point-null hypothesis of 0, you can apply two one-sided tests to boundaries representing your smallest effect size of interest. You apply a one-sided test to a lower bound and see whether your effect is significantly higher. You apply a one-sided test to an upper bound and see whether your effect is significantly lower. If both tests are significant, you can conclude your effect is statistically equivalent, at least within your chosen boundaries. 

You can apply this technique to different focal tests such as t-tests, correlations, and meta-analyses. The package we will use is called `r pkg("TOSTER")`. Make sure you install it before running any of the following code, and load the necessary packages for this appendix. 

```{r load Appendix B packages, message=FALSE, warning=FALSE}
# Applying equivalence testing in R 
library(TOSTER)
# Wrangling and plotting functions
library(tidyverse)
```

To follow along to this appendix and try the code yourself, please download the data files we will be using in [this zip file](data/equivalence_testing.zip).

## t-tests

For a t-test applied to two groups, we will reanalyse data from @bastian_pain_2014. This study wanted to investigate whether experiencing pain together can increase levels of bonding between participants. The study was trying to explain how people often say friendships are strengthened by adversity. 

Participants were randomly allocated into two conditions: pain or control. Participants in the pain group experienced mild pain through a cold pressor task (leaving your hand in ice cold water) and a wall squat (sitting against a wall). The control group completed a different task that did not involve pain. The participants then completed a scale to measure how bonded they felt to other participants in the group. Higher values on this scale mean greater bonding. 

The independent variable is called `r hl("CONDITION")`. The control group has the value 0 and the pain group has the value 1. They wanted to find out whether participants in the pain group would have higher levels of bonding with their fellow participants than participants in the control group. After a little processing, the dependent variable is called `r hl("mean_bonding")` for the mean of 7 items related to bonding. 

```{r Bastian data appendix,  warning=FALSE, message=FALSE}
Bastian_data <- read_csv("data/Bastian.csv")

# Relabel condition to be more intuitive which group is which 
Bastian_data$CONDITION <- factor(Bastian_data$CONDITION, 
                                   levels = c(0, 1), 
                                   labels = c("Control", "Pain"))

# We also need to get our DV from the mean of 7 items
Bastian_data <- Bastian_data %>% 
  pivot_longer(names_to = "item", # var for item names
               values_to = "score", # var for item scores
               cols = group101:group107) %>% # Range of columns for group bonding items
  group_by(across(.cols = c(-item, -score))) %>% # Group by everything but ignore item and score
  summarise(mean_bonding = mean(score)) %>% # Summarise by creating a subscale name and specify sum or mean
  ungroup() # Always ungroup

```

For t-tests, the package has shifted to a single function that can be applied in different ways: 

- For a one-sample t-test, you must enter the mean, SD, and sample size for the sample, alongside the value of mu to test against. 

- For an independent samples t-test, you must enter the mean, SD, and sample size for each group. 

- For a paired samples t-test, you must enter the mean and SD for each condition, the sample size, and the correlation between conditions. 

Bastian et al. requires the independent samples t-test inputs. The first step will be calculating the summary statistics for our two groups and saving them to clearly label and enter into the TOSTER function later. You could enter the values directly, but I like to calculate them down first and save them as objects to avoid copy-paste errors. 

```{r Brandt summary stats}
# Group by condition to get one row for control and pain values
Brandt_summary <- Bastian_data %>% 
  group_by(CONDITION) %>% 
  summarise(mean = mean(mean_bonding),
            sd = sd(mean_bonding),
            n = n()) %>% 
  as.data.frame() # Doesn't work well with tibbles

# Save values for Control
Control_mean <- Brandt_summary[1, 2]
Control_sd <- Brandt_summary[1, 3]
Control_n <- Brandt_summary[1, 4]

# Save values for Pain
Pain_mean <- Brandt_summary[2, 2]
Pain_sd <- Brandt_summary[2, 3]
Pain_n <- Brandt_summary[2, 4]
```

Now we have our values, it is time to think about what values we will use for the boundaries. From the lecture, there are different strategies to choosing and justifying your smallest effect size of interest boundaries: 

1. Your understanding of the applications / mechanisms (e.g., a clinically meaningful decrease in pain). 

2. Smallest effects from previous research (e.g., lower bound of individual study effect sizes or lower bound of a meta-analysis).

3. Small telescopes (the effect size the original study had 33% power to detect).

For Bastian et al. (2014), they measured bonding on a 5-point Likert scale, so we might consider anything less than a one-point difference as too small to be practically meaningful for demonstration purposes. We can now enter these values into the TOSTER function. 

```{r Brandt 1-point TOST}
# Save as object and present with brackets around 
(Brandt_TOST <- tsum_TOST(m1 = Control_mean, # Group 1: Control
                         sd1 = Control_sd,
                         n1 = Control_n,
                         m2 = Pain_mean, # Group 2: Pain
                         sd2 = Pain_sd,
                         n2 = Pain_n,
                         hypothesis = "EQU", # Equivalence test rather than minimal effects test
                         low_eqbound = -1, # Our lower and upper equivalence bounds
                         high_eqbound = 1,
                         eqbound_type = "raw", # We can specify raw or standardised values
                         alpha = 0.05)) # We could change alpha from the default

```

We have quite a lot of output to break down and there are essentially three parts. 

The first provides an overview of the two one-sided test procedure you applied and what you can conclude from it. After confirming your inputs, you have one t-test for the equivalence test and one traditional t-test. You only get one for the equivalence test as both must be significant to conclude the effect was statistically equivalent, so you typically report the largest *p*-value as if that is significant, the other test will be even lower. You then get two one sentence statements on what you can conclude from the tests. For this example, the traditional t-test was significant to suggest we can reject the null hypothesis, but the equivalence test was not statistically equivalent for our choice of boundaries. 

The second provides the full output for all three t-tests: the traditional t-test and the two one-sided tests. 

Finally, the third provides estimates and 90% confidence intervals for the effect size in raw and standardised units. 

For the t-test TOSTER function, we do not get an accompanying plot by default. However, since we saved the model object, we can call a plot to visualise the equivalence testing procedure. 

```{r Brandt TOST plot}
plot(Brandt_TOST)
```


## Correlation

## Meta-analysis

## Further reading

In this appendix, we mainly focused on the practical part of applying equivalence testing in R. Possibly the hardest part of this technique is choosing and justifying your equivalence bounds. Therefore, there are some recommendations for further reading to explore more of the topic. 

- [Lakens (2023)](https://lakens.github.io/statistical_inferences/equivalencetest.html) has an online open access textbook on improving your statistical inferences. Chapter 9 focuses on equivalence testing and explains different methods for setting your smallest effect size of interest. 

- [Lakens (2017)](https://journals.sagepub.com/doi/10.1177/1948550617697177) and [Lakens et al. (2018)](https://journals.sagepub.com/doi/10.1177/2515245918770963) have two tutorial articles where they walk through applying and reporting equivalence testing. 

- [Bartlett et al. (2022)](https://journal.trialanderror.org/pub/attentionalsmoker/release/2) provides a published example of equivalence testing if you want to see how it can look and be reported in a results section. 
